---
title: "EEB177-exercise_8"
output: md_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploring the Cocoli Dataset
```{r}
cocoli_dat <- read.table("/home/eeb177-student/Desktop/eeb-177/class-assignments/classwork-21-Feb/cocoli.txt", header =TRUE)

str(cocoli_dat)
dim(cocoli_dat)
head(cocoli_dat)
```

```{r}

sizes_in_1994 <- cocoli_dat$dbh1
names(sizes_in_1994)<- cocoli_dat$tag
sizes_in_1994[1:10]
```

```{r}
sizes_in_1997 <-cocoli_dat$dbh2
names(sizes_in_1994) <- cocoli_dat$tag

sizes_in_1998 <- cocoli_dat$dbh3
names(sizes_in_1994) <-cocoli_dat$tag
```

```{r}
#find all instaances fo dbh = 171
sizes_in_1994 == 171

which(sizes_in_1994 == 171)

#Let's track fate of trees began 1994 with db 171
trees_171_1994 <- which(sizes_in_1994  == 177)
trees_171_1994

sizes_in_1997[trees_171_1994]

#Calc Relative Growth Rate between 1997 and 1994
yearly_RGR_1 = ((cocoli_dat$dbh2 -cocoli_dat$db1)/cocoli_dat$dbh1)/3

#cocoli_dat$rgr1 <- yearly_RGR_1
#this command doesn't let me add to the dataframe

```


## Plots

```{r}
#make histograms with hsit()
#barplots with barplot()
#boxplot with boxplot()
# scatterplot with l=plot

hist(cocoli_dat$dbh1, xlim = c(0,2000), xlab="DBH in 1994 (mm)", main ="Distribution of sizes in 1994", col="red")

help(hist)
```

## Packages
```{r}
library(ggplot2)
ggplot(cocoli_dat) + geom_histogram(aes(dbh1), fill ="blue", alpha =0.2) + geom_histogram(aes(dbh2), fill = "red", alpha= 0.2) + ggtitle("Distribution of sizes in 1994")
```


## In Class assignments

```{r}
housing <- read.csv("Rgraphics/dataSets/landdata-states.csv")
head(housing[1:5])
```
```{r}
hist(housing$Home.Value)
#just charts a histogram
```

```{r}
library(ggplot2)
ggplot(housing, aes(x = Home.Value)) + #specifies the data, and the x axis
  geom_histogram()
#then just says to make a histogram
```


```{r}
plot(Home.Value ~ Date,
     data=subset(housing, State == "MA"))
#yasy plot a subset of data, for Mass
points(Home.Value ~ Date, col="red",
       data=subset(housing, State == "TX"))
#says put points on existing graph from subset of points from Texas
legend(1975, 400000,
       c("MA", "TX"), title="State",
       col=c("black", "red"),
       pch=c(1, 1))
#says make a legend with title "State", and the colors black is MA, red is TX
```

```{r}
ggplot(subset(housing, State %in% c("MA", "TX")),
       aes(x=Date,
           y=Home.Value,
           color=State))+
  #wants a plot using the subset of data from housing, with MA and TX, x axis as date, y value as home value, automatically assigning color by state
  geom_point()
#want to plot by points?
```

## Points (Scatterplot)
```{r}
hp2001Q1 <- subset(housing, Date == 2001.25) 
ggplot(hp2001Q1,
       aes(y = Structure.Cost, x = Land.Value)) +
  geom_point()
#geom points requires mapping for x and y
# y is Structure cost
# x is land value
```

```{r}
ggplot(hp2001Q1,
       aes(y = Structure.Cost, x = log(Land.Value))) +
  geom_point()
#we've switched the x and y axes for this graph
```

## Lines (Predicion Lines)
When I Knit it, his next line doesn't work for some reason, can't assign anything to dataset
messes up the whole next section,
worked when I first tried it
removing the chunks to make the doc actually knit

hp2001Q1$pred.SC <- predict(lm(Structure.Cost ~ log(Land.Value), data = hp2001Q1))
#adding a column to the dataset, cost-value

p1 <- ggplot(hp2001Q1, aes(x = log(Land.Value), y = Structure.Cost))
# creating p1, from the dataset, with x as land, and y as structurecost,p1 +
  geom_point(aes(color = Home.Value)) +
  geom_smooth()

p1 + geom_point(aes(color = Home.Value)) +
  geom_line(aes(y = pred.SC))
#geometric points, assign color based on home value
# assigning the prediction line to the scatterplot



## Smoothers
```{r}
p1 <- ggplot(hp2001Q1, aes(x = log(Land.Value), y = Structure.Cost))
p1 +
  geom_point(aes(color = Home.Value)) +
  geom_smooth()
# making a smooth line, rather than a straight one
```

## Text (Label Points)
```{r}
p1 + 
  geom_text(aes(label=State), size = 3)
#same thing as above, no prediction line,
#labels it according to state
```
```{r}
## install.packages("ggrepel") 
library("ggrepel")
p1 + 
  geom_point() + 
  geom_text_repel(aes(label=State), size = 3)
#uses ggrepel to make the labels not crowd eachother
#also gives the dot and the label, not just the label
```

```{r}
#filesare mapped to aesthetics with aes() function, whereas fixed aesthetics are set outside the aes() call

p1 +
  geom_point(aes(size = 2),# incorrect! 2 is not a variable
             color="red") # this is fine -- all points red

```


## Mapping variables to other aesthetics

```{r}
p1 +
  geom_point(aes(color=Home.Value, shape = region))
#map other aesthetics the same way as x and y in the previous example
```

## Exercise I
#### I.1 scatterplot with CPI on x axis, HDI on y axis
```{r}
dat <- read.csv("Rgraphics/dataSets/EconomistData.csv")

ggplot(dat, aes(x = CPI, y = HDI)) + geom_point()
```
####I.2 color the points blue
```{r}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point(color = "blue")
```
##### I.3 map the color to the region
```{r}
ggplot(dat, aes(x = CPI, y = HDI, color)) + geom_point()
```
```{r}
unique(dat$Region)
```


##### I.4 make points bigger by setting size to 2
```{r}
ggplot(dat, aes(x = CPI, y = HDI, color = Region)) + geom_point(size=2)
```
##### I.5 Map point to HDI.Rank
```{r}
ggplot(dat, aes(x = CPI, y = HDI, color = Region, size = HDI.Rank)) + geom_point()
```
```{r}
args(geom_histogram)
args(stat_bin)
```
```{r}
p2 <- ggplot(housing, aes(x = Home.Value))
p2 + geom_histogram()
```
```{r}
p2 + geom_histogram(stat = "bin", binwidth = 4000)
```
```{r}
housing.sum <- aggregate(housing["Home.Value"], housing["State"], FUN=mean)
rbind(head(housing.sum), tail(housing.sum))
```
```{r}
ggplot(housing.sum, aes(x=State, y=Home.Value)) + geom_bar(stat ="identity")
```
## Exercise II
II.
```{r}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point()
```
II.2
```{r}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point() + geom_smooth()
```
II.3
```{r}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point() + geom_smooth(method = lm)
```
II.4
```{r}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point() + geom_line(stat = "ecdf")
```
```{r}
p3 <- ggplot(housing,
             aes(x = State,
                 y = Home.Price.Index)) + 
        theme(legend.position="top",
              axis.text=element_text(size = 6))
(p4 <- p3 + geom_point(aes(color = Date),
                       alpha = 0.5,
                       size = 1.5,
                       position = position_jitter(width = 0.25, height = 0)))
#makes the p4 object which is p3 plus the points passed, makes color a scale according to Date
#puts in position jitter for somereason

```
```{r}
p4 + scale_x_discrete(name="State Abbreviation") +
  scale_color_continuous(name="",
                         breaks = c(1976, 1994, 2013),
                         labels = c("'76", "'94", "'13"))
#modifies the key to be broken up by the appropriate numbers
```
```{r}
p4 +
  scale_x_discrete(name="State Abbreviation") +
  scale_color_continuous(name="",
                         breaks = c(1976, 1994, 2013),
                         labels = c("'76", "'94", "'13"),
                         low = "blue", high = "red")
```
```{r}
p4 +
  scale_color_gradient2(name="",
                        breaks = c(1976, 1994, 2013),
                        labels = c("'76", "'94", "'13"),
                        low = "blue",
                        high = "red",
                        mid = "gray60",
                        midpoint = 1994)
#note that muted() doesn't seem to exist so I got rid of it
```
## Exercise III
III.1
```{r}
ExPlot <- ggplot(dat, aes(x = CPI, y = HDI), scale_y_continuous("Human Development Index")) + geom_point(aes(color= Region))
ExPlot
```
III.2
```{r}
Explot2 <- ExPlot + scale_x_continuous(name = "Corruption perception index") + scale_y_continuous(name= "Human development index") + scale_fill_manual(name= "Region", breaks= c("Americas", "Asia Pacific", "East EU Cemt Asia", "EU W. Europe", "MENA", "SSA"), labels = c("1", "2", "3", "4", "5", "6"))
#this should relabel the labels but does not for some reason, I've tried this several times
Explot2
```

```{r}
?scale_color_manual
ExPlot <- ggplot(dat, aes(x = CPI, y = HDI), scale_y_continuous("Human Development Index")) + geom_point(aes(color= Region))
cols <- c("deepskyblue1", "cyan", "cyan4", "dodgerblue4", "red1", "firebrick4")
Explot3 <- Explot2 + scale_color_manual(values = cols)
Explot3
```
```{r}
p5 <- ggplot(housing, aes(x = Date, y = Home.Value))
p5 + geom_line(aes(color = State))
```
```{r}
(p5 <- p5 + geom_line() +
   facet_wrap(~State, ncol = 10))
```

```{r}
p5 + theme_linedraw()
```
```{r}
p5 + theme_linedraw()
```
```{r}
p5 + theme_minimal() +
  theme(text = element_text(color = "turquoise"))
```

```{r}
theme_new <- theme_bw() +
  theme(plot.background = element_rect(size = 1, color = "blue", fill = "black"),
        text=element_text(size = 12, family = "Serif", color = "ivory"),
        axis.text.y = element_text(colour = "purple"),
        axis.text.x = element_text(colour = "red"),
        panel.background = element_rect(fill = "pink"),
        strip.background = element_rect(fill = "orange"))

p5 + theme_new
```

```{r}
#wrong way to plot diff varaibles with diff colors
housing.byyear <- aggregate(cbind(Home.Value, Land.Value) ~ Date, data = housing, mean)
ggplot(housing.byyear,
       aes(x=Date)) +
  geom_line(aes(y=Home.Value), color="red") +
  geom_line(aes(y=Land.Value), color="blue")

```

```{r}
#right way to do it
library(tidyr)
home.land.byyear <- gather(housing.byyear,
                           value = "value",
                           key = "type",
                           Home.Value, Land.Value)
ggplot(home.land.byyear,
       aes(x=Date,
           y=value,
           color=type)) +
  geom_line()
```

#Challenge to recreate Econ Graph
```{r}
dat <- read.csv("Rgraphics/dataSets/EconomistData.csv")
#puts a red trendline on the graph
Econgraph <- ggplot(dat, aes(x = CPI, y = HDI)) + geom_point() + geom_smooth(method = "lm", formula = y~ poly(x,2), color="red")
```
```{r}
# let's add axis labels
Econgraph2 <- Econgraph + scale_x_continuous(breaks =seq(0,10) , name = "Corruption perception index") + scale_y_continuous(name= "Human development index")

```

```{r}
#add some color
cols <- c("deepskyblue1", "cyan", "cyan4", "dodgerblue4", "red1", "firebrick4")
#defines the color vector
Econgraph3 <-Econgraph2 + geom_point(aes(color= Region, shape =1), shape = 1, fill=NA) + scale_color_manual(values = cols) +theme_bw() + theme(legend.position = "top", legend.direction = "horizontal", panel.grid.major.x =  element_blank(), panel.grid.minor.x = element_blank())
#added color
#put legend on top of the grid
# changed the shape of the object

```
```{r}
library("ggrepel")

Nations_of_interest <- c("Congo","Afghanistan", "Sudan", "Myanmar", "Iraq", "Venezula", "Russia", "India", "Argentina", "Greece", "Brazil", "Italy", "China", "South Africa", "Rwanda", "Bhutan", "Cape Verde", "Botswana", "Spain", "France", "US", "Britain","Barbados", "Germany", "Japan", "Norway", "Singapore", "New Zealand")

#then need to subset the data based on the presence of these words in the Country category
subset_of_interest <- dat[dat$Country %in% Nations_of_interest,]

```

```{r}
Econgraph4 <- Econgraph3 + geom_text_repel(data = subset_of_interest, aes(label=Country), size =3)+ scale_fill_manual(name= "Region", breaks= c("Americas", "Asia Pacific", "East EU Cemt Asia", "EU W. Europe", "MENA", "SSA"), labels = c("OECD", "Americas", "Asia & Oceania", "Central & Eastern Europe", "Middle East & north Africa", "Sub-Saharan Africa"))

#for somereason it gives the correct result but quickly boots it off afterwards, it si aggravating
#also the breaks and names don't work
Econgraph4
```
```{r}
```


